### 2017-11-02 paper/abstract.tex
backu

\begin{abstract}

In recent years, The end of Dennard Scaling \cite{Dark-Silicon} and the need for agile development methodology in data centers is pushing the computer architecture community towards designing more specialized and energy efficient systems \cite{Catapult}. This has lead to a new application domain for FPGAs and the emergence of heterogeneous CPU-FPGA computing platforms \cite{Intel-FPGA} \cite{IBM-FPGA}, enabling the design of new energy efficient FPGA accelerators for domain specific applications. Graph Analytics, one the largest applications in production data centers today, faces scaling challenges with the ever increasing size of its workloads and its inherent low locality and memory-bound characteristic. The GraphMat \cite{GraphMat} programming framework was proposed that efficiently translates graph algorithms into high performance computation in hardware. We present FlexGraph, a reconfigurable domain-specific accelerator for Graph Analytics targeting commodity heterogeneous CPU-FPGA platforms. FlexGraph is designed using Cocoh \cite{Cocoh}, a C++ domain specific library for hardware design and simulation, providing a high level programming abstraction with the efficiency of native RTL. FlexGraph's novel architecture allows an easy implementation of GraphMat kernels using standard C++. To the best of our knowledge it is a first implementation of a graph accelerator on a commodity heterogeneous CPU-FPGA platform. FlexGraphs's preliminary performance is comparable to HLS, we later discuss several optimization strategies that could further improve its speed.

\end{abstract}