### 2017-11-02 paper/evaluation.tex

\section{Experimental Setup}

\subsection{Cocoh's Simulation}

\subsection{Intel AAL Interface}

\subsection{Intel QPI Emulation}

\subsection{Graph Datasets}

\section{Results Analysis}

\subsection{Overall Performance}

\subsection{The Impact of Parallelism}

\subsection{The Impact of Memory Optimizations}

\subsection{FlexGraph Performance Hogs}

\subsection{Hardware Cost Evaluation}