### 2017-11-15 paper/abstract.tex
backup

\begin{abstract}

In recent years, The end of Dennard Scaling is pushing the computer architecture community towards designing more specialized and energy efficient systems. This move has led to a new application domain for FPGAs and the emergence of heterogeneous CPU-FPGA computing platforms, enabling the design of new energy efficient FPGA accelerators for domain specific applications. Graph Analytics, one the largest applications in production data centers today, faces scaling challenges with its ever increasing workload size, inherent sparsity and memory-bound characteristic. We present FlexGraph, a flexible and energy-efficient Graph Analytics Accelerator for heterogeneous CPU-FPGA architectures. FlexGraph uses a Doubly Compressed Sparse Matrix format to eliminate unnecessary data transfers and reduce storage requirements. Our architecture allows support for other compressed format by decoupling the matrix data structure traversal from the compute and memory units. We introduce memory access hardware primitives for unstructured fine-grained accesses on cache-coherent shared memory. Flexgraph uses a C++ Hardware Generation Language to extend its vertex programming model with domain centric reconfigurability in an integrated single source development environment, providing a design efficient alternative to High Level Synthesis. Flexgraph achieves X GFlop/s, performing Nx faster than HLS implementation.

\end{abstract}